Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-10-12 experiments.tex
AA: fornendo motivazioni alle metriche scelte

% ACM table style
\begin{table}
	\tbl{Benchmarks\label{tab:bench}}{
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
	\hline
	Alloca\_test & Memory array allocation test \\
	Bmm & Matrices multiplication and elements sum \\
	Fir\_int & Finite impulse response \\
	Mm & Floating point matrices multiplication \\
	Sha256 & Cryptocurrency header Hashing \\
	Wave & Wavefront computation \\
	\hline
	\end{tabular}}
\end{table}
\subsection{Evaluation Metrics}

@@

@@

In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. For a given
objective, the variation range represents the ratio between the
maximum and the minimum value observed for that objective.
A comparison between the variation ranges of both power dissipation
and execution time, obtained by PS and GA for different benchmarks is
shown in \figR{dispersion}(a) and (b) respectively.  As it can
be observed, the PS exploration provides solutions which fall on a
range that is, on average, wider than the one provided by a GA
exploration for power dissipation and execution time, respectively.
%\begin{table}
%  \centering
%  \begin{tabular}{c}
%    \includegraphics[width=0.7\textwidth]{pictures/variation_power.eps} \\
%    (a) \\
%    \includegraphics[width=0.7\textwidth]{pictures/variation_etime.eps} \\
%    (b) \\
%    \includegraphics[width=0.7\textwidth]{pictures/dispersion.eps} \\
%    (c) 
%  \end{tabular}
%  \caption{(a)(b) Variation range obtained by PS and GA for different benchmarks
%  and average dispersion absolute dispersion (c)}
%  \label{fig:dispersion}
%\end{table}

@@

The average normalised absolute dispersion error measures
the average absolute difference between the distribution of points in
the objective space and an ideal distribution in which the points are
uniformly distributed over the objective space. Formally, let $O$ be
the image, in the objective space, of the configurations visited by
the design space exploration. The generic element of $O$ (\ie, a
solution) is a pair $(p,t)$ where $p$ and $t$ are the average power
and execution time, respectively. The two-dimensional objective space
is then partitioned by a $M_x \times M_y$ mesh. For each tile $T_i,
\ i=1, 2, \ldots, M_xM_y$ of the mesh, let $N_i$ be the number
of points in $O$ which fall in $T_i$. The average absolute error, $E_i$, for
$T_i$ is the absolute value of the difference between $N_i$ and the
ideal number of solutions, $\overline{N}$, which should fall in $T_i$
in case of uniform distribution. Such $\overline{N}$ can be simply
computed as the ratio between the cardinality of $O$ and the number of
tiles. Thus,
\[ E_i = |N_i - \overline{N}|, \]
where $\overline{N} = |O| / (M_x M_y)$. The average
normalised absolute dispersion error ($ANADE$) is the average of $E_i$
normalised to the maximum absolute error $E_{max}$:
\[ ANADE = \frac{\sum_{i=1}^{M_xM_y} E_i/(M_xM_y)}{E_{max}}, \]
where $E_{max}$ can be computed as the average absolute error in the worst
case in which all the solutions fall in a single tile:
\[ E_{max} = \frac{(M_x M_y - 1) \overline{N} + |\overline{N} -
    |O|| }{M_x M_y}. \]

>>>>

% ACM table style
\begin{table}
	\tbl{Benchmarks\label{tab:bench}}{
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
	\hline
	Alloca\_test & Memory array allocation test \\
	Bmm & Matrices multiplication and elements sum \\
	Fir\_int & Finite impulse response \\
	Mm & Floating point matrices multiplication \\
	Sha256 & Cryptocurrency header Hashing \\
	Wave & Wavefront computation \\
	\hline
	\end{tabular}}
\end{table}
\subsection{Terms of comparison}

@@

\subsection{Evaluation metrics}
In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}.

For a given objective, the \emph{variation range} represents the ratio between the maximum and the minimum value observed for that objective.
\comment{AA: Per rappresentare l'extent del Pereto front, forse sarebbe meglio evitare di usare il variation range, per come l'hai definito tu, e usare la funzione M3 di \cite{zitzler_ec00}, Sec.5 ``Metrics of performance'', che mi pare rappresenti la stessa cosa, concettualmente. Questo ci permette di essere meno attaccabili (del tipo ``non te la prendere con noi, prenditela con Zitzler se non capisci l'utilita' della metrica'').}
It quantifies the extent of the of the Pareto front that is an important factor to consider when evaluating the quality of a Pareto front. Indeed, authors in \cite{zitzler_ec00} carefully observe that an exploration algorithm should preferably provide pareto fronts whose objective function range is large.
Such Pareto fronts permit to investigate in a broader way the potentials of the designed device. This is important, in general, and turns out to be essential when designing a device that may be required to work in different operational conditions, for example with stringent execution time constraints and less strict power dissipation ones or, on the contrary, with low power requirements and tolerance to high execution time.

The average \emph{normalised absolute dispersion error} measures
the average absolute difference between the distribution of points in
the objective space and an ideal distribution in which the points are
uniformly distributed over the objective space. 
The motivation behind this metric is, again, given by \cite{zitzler_ec00}, in which authors remark that a desirable property of the Pareto front is the spread of the non-dominated configurations all over the configuration space. In other words, the more similar the non-dominated configuration distribution is to the uniform distribution, the better is the quality of the Pareto front.
\comment{AA: commentare perche' e' importante questa metrica. Cercare la parola chiave ``deception'' o qualcosa del genere}.
Formally, let $O$ be
the image, in the objective space, of the configurations visited by
the design space exploration. The generic element of $O$ (\ie, a
solution) is a pair $(p,t)$ where $p$ and $t$ are the average power
and execution time, respectively. The two-dimensional objective space
is then partitioned by a $M_x \times M_y$ mesh. For each tile $T_i,
\ i=1, 2, \ldots, M_xM_y$ of the mesh, let $N_i$ be the number
of points in $O$ which fall in $T_i$. The average absolute error, $E_i$, for
$T_i$ is the absolute value of the difference between $N_i$ and the
ideal number of solutions, $\overline{N}$, which should fall in $T_i$
in case of uniform distribution. Such $\overline{N}$ can be simply
computed as the ratio between the cardinality of $O$ and the number of
tiles. Thus,
\[ E_i = |N_i - \overline{N}|, \]
where $\overline{N} = |O| / (M_x M_y)$. The average
normalised absolute dispersion error ($ANADE$) is the average of $E_i$
normalised to the maximum absolute error $E_{max}$:
\[ ANADE = \frac{\sum_{i=1}^{M_xM_y} E_i/(M_xM_y)}{E_{max}}, \]
where $E_{max}$ can be computed as the average absolute error in the worst
case in which all the solutions fall in a single tile:
\[ E_{max} = \frac{(M_x M_y - 1) \overline{N} + |\overline{N} -
    |O|| }{M_x M_y}. \]

@@

A comparison between the variation ranges of both power dissipation
and execution time, obtained by PS and GA for different benchmarks is
shown in \figR{dispersion}(a) and (b) respectively.  As it can
be observed, the PS exploration provides solutions which fall on a
range that is, on average, wider than the one provided by a GA
exploration for power dissipation and execution time, respectively.
%\begin{table}
%  \centering
%  \begin{tabular}{c}
%    \includegraphics[width=0.7\textwidth]{pictures/variation_power.eps} \\
%    (a) \\
%    \includegraphics[width=0.7\textwidth]{pictures/variation_etime.eps} \\
%    (b) \\
%    \includegraphics[width=0.7\textwidth]{pictures/dispersion.eps} \\
%    (c) 
%  \end{tabular}
%  \caption{(a)(b) Variation range obtained by PS and GA for different benchmarks
%  and average dispersion absolute dispersion (c)}
%  \label{fig:dispersion}
%\end{table}

@@

### 2014-10-12 introduction.tex
AA: fornendo motivazioni alle metriche scelte

In this work we present PS, a multiobjective design exploration
algorithm that introduces the concept of \emph{interesting} and
\emph{uninteresting regions} of the configuration space.  The main
idea can be simplified as follows: at each iteration of the algorithm,
the configuration space is divided in regions. A score is given to
each region based on the \emph{innovation} provided by its
configurations in the Pareto front. The regions with high score are
considered the most \emph{interesting} and more simulations are
distributed there rather than in the rest of the configuration space.
Several geometrical transformations (splitting, merging etc.) are
involved in the Parameter Space (aka configuration space), hence
the name PS for the algorithm.

@@

Many studies, as \cite{coello_easmop} and others, focus on genetic approaches to solve multiobjective
optimization problems. Genetic approaches
have many advantages: they provide good accuracy, they are customizable and very general
and require no a-priori knowledge to the designer.  The strong
point of genetic approaches can be summarized saying that they consist
in exploration that is sufficiently broad (the mutation avoids being rigidly restricted to limited parts of parameter space), and, at the same time, not too scattered as it is guided by the performances of
the already evaluated configurations. In this work we will compare 
the proposed approach against a well known Multi-Objective Genetic
Algorithm~\cite{knowles_techrep06}.
We claim that the approach presented in this paper has most of the benefits of the genetic
approaches, although the rationale is completely different.

>>>>

In this work we present PS, a multiobjective design exploration
algorithm that introduces the concept of \emph{interesting} and
\emph{uninteresting regions} of the configuration space.  The main
idea can be simplified as follows: at each iteration of the algorithm,
the configuration space is divided in regions. A score is given to
each region based on the \emph{innovation} provided by its
configurations in the Pareto front, i.e. how different these configurations are, in terms of objectives, with respect to the ones already found previously. The regions with high score are considered the most \emph{interesting} and more simulations are distributed there rather than in the rest of the configuration space.
Several geometrical transformations (splitting, merging etc.) are
involved in the Parameter Space (aka configuration space), hence
the name PS for the algorithm.

@@

\cite{dellnitz2005covering} proposes three different methods for the numerical approximation of the Pareto set. The first two require the knowledge of the gradient of the objective functions, which is usually not available in embedded system design, since the closed form expression of the objective functions is not known a priori. The third one, called ``Sampling algorithm'', does not need this requirement and is the closest to ours, although the mechanism is different. At every iteration a set of regions of configuration space is constructed with the aim to cover the Pareto set. The size of these regions becomes smaller at each iteration so that the covering becomes tighter. A random set of test points are evaluated in each new region; only the regions containing non-dominated points will be considered in the next iteration. Our approach is different. First, in the Sampling algorithm only the selected regions are examined, even though there is no guarantee that all the Pareto points are included in them. As a consequence, the Sampling algorithm may erroneously and prematurely neglect some regions that may possibly contain a non-negligible number of Pareto points. Our algorithm avoids this, since even though a region has not provided non-dominated points, it is not neglected, rather it is simply populated with fewer test points, thus giving the chance to find non-dominated points in them. From this behavior also comes another important difference. The Sampling algorithm of \cite{dellnitz2005covering} keeps reinforcing the analysis of the regions selected in the previous iterations. In other words, it keeps zooming into the regions that have been considered ``interesting'', i.e. rich of non-dominated points, in the previous iterations. The philosophy behind our algorithm is different. A region, which has been considered interesting at a certain iterations, may be not in the future, since it has been thoroughly analyzed. On the other hand, a region, initially uninteresting, may turn out to be interesting in the future, once all the other regions have been already exploited. Therefore, from an iteration to the next, our algorithm zoom in on regions that are currently interesting and zoom out from the uninteresting ones,  spanning in a broader and more dynamical way the configuration space. In addition, in the Sampling region only the number of non-dominated points is accounted to see how interesting a region is. On the contrary, we also consider the \emph{innovation}, as already anticipated in the introduction.

Many studies, as \cite{coello_easmop} and others, focus on genetic approaches to solve multiobjective
optimization problems. Genetic approaches
have many advantages: they provide good accuracy, they are customizable and very general
and require no a-priori knowledge to the designer.  The strong
point of genetic approaches can be summarized as follows: they consist
in an exploration that is sufficiently broad (the mutation avoids being rigidly restricted to limited parts of parameter space), and, at the same time, not too scattered as it is guided by the performances of
the already evaluated configurations. In this work we will compare 
the proposed approach against a well known Multi-Objective Genetic
Algorithm~\cite{knowles_techrep06}.
We claim that the approach presented in this paper has most of the benefits of the genetic
approaches, although the rationale is completely different.
